Introductory Physics Labs: A Tale of Two Transformations

### 2021-03-26 manuscript/rev1docs/response.tex

\documentclass{letter}

\usepackage[top=1in,left=1.5in,right=1.5in]{geometry}

\makeatletter
\renewcommand{\closing}[1]{\par\nobreak\vspace{\parskip}%
  \stopbreaks
  \noindent
  \ifx\@empty\fromaddress\else
  \hspace*{\longindentation}\fi
  \parbox{\indentedwidth}{\raggedright
       \ignorespaces #1\\[2\medskipamount]%
       \ifx\@empty\fromsig
           \fromname
       \else \fromsig \fi\strut}%
   \par}
\makeatother

\signature{The Authors of \textit{Introductory Physics Labs: A Tale of Two Transformations}}

\usepackage{xcolor}
\newcommand{\swsays}[1]{\textcolor{red}{SW says: #1}}

\begin{document}
\begin{letter}{Dr.\ Gary White \\ Editor: \textit{The Physics Teacher} \\ American Association
    of Physics Teachers \\ College Park, MD}

  \opening{Dear Dr.\ White and referee,}

  Thank you for your thoughtful review of our article entitled \textit{Introductory Physics
    Labs: A Tale of Two Transformations}. We have considered your review and are responding to
  your feedback in the following ways.

  \begin{description}
    \item[Institutional support for students needs] Dr.\ White wanted to know more about the
    ways that our institution attempted to support students during the pandemic.  Our office of
    student affairs developed programs and outreach to help support students such as
    disseminating CARES Act funding to students demonstrating exceptional need.  We do discuss
    academic measures taken by the university such as extending the withdrawal date, exempting
    withdrawals during the Spring 2020 and Fall 2020 semesters from the overall limit on
    withdrawals, and allowing students to be graded pass/fail after seeing their final grade
    (and having that grade not count toward their GPA).
    \item[Comparing ADI to Modeling Instruction] We've added a few sentences to the
    introduction comparing ADI to Modeling Instruction.  We also discuss ISLE (Investigative
    Science Learning Environment) as it is also a similar pedagogical approach driven by
    learning cycles.  Both Modeling and ISLE tend to be implemented in lecture courses, while
    ADI was designed to mimic the ways that scientists learn in the laboratory, and has been
    designed to be implemented in laboratory courses regardless of the approach used in
    lecture. We find it to be particularly useful in our context as students in each of our lab
    courses take two different intro physics lecture courses (algebra-based and calculus-based).
    \item[Clarity about proposals and proposal approvals] We've added a few sentences talking
    about how we generated datasets during Spring 2020 and approve proposals in general. We
    want to point out that we do allow some proposals to go forward even if they have a flaw
    once data is collected.  Students are free to revise their proposals and gather more
    data. Dealing with unexpected results can be a powerful teacher. As far as creating
    datasets goes, we tried to make these data as raw as possible, for example sharing videos
    of a object oscillating on a spring and the mass of that object rather than giving them
    mass and period data directly. \swsays{Mark, do you want to say something about how you
      chose what data to generate?}
    \item[Details about Fall 2020] For the Fall 2020 semester, all data were collected by the
    students.  This includes images and videos as appropriate, often collected with students'
    smartphones. \swsays{Mark, please fact-check the above statement.}
    \item[Details about the survey] The survey was conducted during the spring/early summer of
    2020.  Results were released to the faculty before the fall 2020 semester began.  This was
    intended to help the faculty better understand the difficulties that were impacting our
    students. The date of the survey has been added to the text in the conclusion.
  \end{description}

  Thank you for your time.

  \closing{Sincerely,}

\end{letter}

\end{document}

### 2021-03-30 manuscript/rev1docs/response.tex

\begin{document}
\begin{letter}{Dr.\ Gary White \\ Editor: \textit{The Physics Teacher} \\ American Association
    of Physics Teachers \\ College Park, MD}

  \opening{Dear Dr.\ White and referee,}

  Thank you for your thoughtful review of our article entitled \textit{Introductory Physics
    Labs: A Tale of Two Transformations}. We have considered your review and are responding to
  your feedback in the following ways.

  \begin{description}
    \item[Institutional support for students needs] Dr.\ White wanted to know more about the
    ways that our institution attempted to support students during the pandemic.  Our office of
    student affairs developed programs and outreach to help support students such as
    disseminating CARES Act funding to students demonstrating need.  We do discuss academic
    measures taken by the university such as extending the withdrawal date, exempting
    withdrawals during the Spring 2020 and Fall 2020 semesters from the overall limit on
    withdrawals, and allowing students to be graded pass/fail after seeing their final grade
    (and having that grade not count toward their GPA).
    \item[Comparing ADI to Modeling Instruction] We've added a few sentences to the
    introduction comparing ADI to Modeling Instruction.  We also discuss ISLE (Investigative
    Science Learning Environment) as it is also a similar pedagogical approach driven by
    learning cycles.  Both Modeling and ISLE tend to be implemented in lecture courses, while
    ADI was designed to mimic the ways that scientists learn in the laboratory, and has been
    designed to be implemented in laboratory courses regardless of the approach used in
    lecture. We find it to be particularly useful in our context as students in each of our lab
    courses take two different intro physics lecture courses (algebra-based and calculus-based).
    \item[Clarity about proposals and proposal approvals] We've added a few sentences talking
    about how we generated datasets during Spring 2020 and approve proposals in general. We
    want to point out that we do allow some proposals to go forward even if they have a flaw
    once data is collected.  Students are free to revise their proposals and gather more
    data. Dealing with unexpected results can be a powerful teacher. As far as creating
    datasets goes, we tried to make these data as raw as possible, for example sharing videos
    of a object oscillating on a spring and the mass of that object rather than giving them
    mass and period data directly. 
    \item[Details about Fall 2020] For the Fall 2020 semester, all data were collected by the
    students.  This includes images and videos as appropriate, often collected with students'
    smartphones. Additional text has been added to Section \textrm{IV}.B.
    \item[Details about the survey] The survey was conducted during the spring/early summer of
    2020.  Results were released to the faculty before the fall 2020 semester began.  This was
    intended to help the faculty better understand the difficulties that were impacting our
    students. The date of the survey has been added to the text in the conclusion.
  \end{description}

### 2021-03-30 manuscript/onlineLabPaper.tex
Updated the manuscript including all changes indicated in the response letter.

The physics labs at our institution in Fall 2016 very much mirrored the national narrative.
The lab curriculum had been largely unchanged for about 20 years. Moreover, the equipment and
techniques used by students was outdated---triple beam balances and hand-drawn graphs were
standard.  That same year, as part of an interdisciplinary team interested in engaging students
in and assessing scientific practices, we secured funding, both internally and from the NSF, to
change this \emph{status quo} in physics, biology, and chemistry.  All three departments
transformed their labs using the Argument Driven Inquiry
(ADI)\citep{Sampson2011,Walker2011,Walker2016} pedagogical framework, which has been
specifically designed to engage students in authentic science practices. In physics, we also
utilized the AAPT Lab Recommendations to guide our transformation.\citep{kozminski2014aapt}

We were satisfied with our progress when the global COVID-19 pandemic forced a shutdown in
March 2020.  With students off campus, we had to abruptly shift our delivery to an online
format and quickly realized that many of the face-to-face activities had to be adapted to
online use.  Going forward, we have further adapted what we are doing so that students can
still participate in a hands-on laboratory experience in an online laboratory course.

@@

The transition to online learning has not gone without hiccups.  Technology and access are two
key factors for success in online learning environments, which were a barrier for many of our
students.  Our institution's Center for Survey Research has published some findings that
explain some of our observed issues.\citep{ECUcovidSurveyRedacted} The GTAs and students found
communication about proposals much more difficult online than in face-to-face classes.  About
\SI{10}{\percent} to \SI{20}{\percent} of our students did not have reliable internet access
after the transition to online learning in the spring.\citep{ECUcovidSurveyRedacted} Also, up
to \SI{8}{\percent} of students did not have access to a computer once they left
campus.\citep{ECUcovidSurveyRedacted} Additionally, we noticed that some students in the course
did not have access to computers capable of running \emph{Tracker} or \emph{ImageJ}, both of
which run on Windows, Macintosh, or Linux computers but not Chromebooks or mobile devices. We
discovered \emph{jsTrack}\citep{jstrack}, an online Javascript web application for video
analysis that runs on most computers including Chromebooks, but not mobile devices.

>>>>

The physics labs at our institution in Fall 2016 very much mirrored the national narrative.
The lab curriculum had been largely unchanged for about 20 years. Moreover, the equipment and
techniques used by students was outdated---triple beam balances and hand-drawn graphs were
standard.  That same year, as part of an interdisciplinary team interested in engaging students
in and assessing scientific practices, we secured funding, both internally and from the NSF, to
change this \emph{status quo} in physics, biology, and chemistry.  All three departments
transformed their labs using the Argument Driven Inquiry
(ADI)\citep{Sampson2011,Walker2011,Walker2016} pedagogical framework, which has been
specifically designed to engage students in authentic science practices. ADI is very much in
the spirit of transformations such as Modeling Physics\cite{modelingBrewe2008} and the
Investigative Science Learning Environment (ISLE) \cite{etkina2007investigative}. Oftentimes,
both Modeling Physics and ISLE are implemented in studio-style or even lecture classrooms,
whereas ADI is lab-focused. The common use of hybrid courses for Modeling and ISLE imply a
shorter time scale, for example, an ISLE cycle should take place over the course of a week,
whereas the cycles that we use in ADI take place over 4 weeks. For our institutional
transformation of our physics labs with ADI, we also utilized the AAPT Lab Recommendations to
inform our curricular choices.\citep{kozminski2014aapt}

We were satisfied with our progress when the global COVID-19 pandemic forced a shutdown in
March 2020.  With students off campus, we had to abruptly shift our delivery to an online
format and quickly realized that many of the face-to-face activities had to be adapted to
online use.  Going forward to Fall 2020 (and beyond), we have further adapted what we are doing
so that students can still participate in a hands-on laboratory experience in an online
laboratory course.

@@

The transition to online learning has not gone without hiccups.  Technology and access are two
key factors for success in online learning environments, which were a barrier for many of our
students.  Our institution implemented a number of academic measures intended to support our
students during the pandemic, including extending the withdrawal date and allowing students to
be graded pass/fail after seeing their final grade during each of the Spring 2020 and Fall 2020
semesters. There have also been technical supports put in place at the institution such as
extended laptop loan programs and remote proctoring was paid for by a grant through the
university system. Additionally, our institution's Center for Survey Research has published
some findings in August 2020 that explain some of our observed
issues.\citep{ECUcovidSurveyRedacted} The GTAs and students found communication about proposals
much more difficult online than in face-to-face classes.  About \SI{10}{\percent} to
\SI{20}{\percent} of our students did not have reliable internet access after the transition to
online learning in the spring.\citep{ECUcovidSurveyRedacted} Also, up to \SI{8}{\percent} of
students did not have access to a computer once they left campus.\citep{ECUcovidSurveyRedacted}
Additionally, we noticed that some students in the course did not have access to computers
capable of running \emph{Tracker} or \emph{ImageJ}, both of which run on Windows, Macintosh, or
Linux computers but not Chromebooks or mobile devices. We discovered
\emph{jsTrack}\citep{jstrack}, an online Javascript web application for video analysis that
runs on most computers including Chromebooks, but not mobile devices.
